---
title: "Statistics Warmup"
author: "Yim Register"
date: "8/21/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Learning to Code

https://github.com/freeCodeCamp/2016-new-coder-survey#about-the-data

This is a survey of over 15,000 respondants about their experiences of learning to code. You can read more about their goals, process, and survey [here](https://www.freecodecamp.org/news/we-just-launched-the-biggest-ever-survey-of-people-learning-to-code-cac81dadf1ea/#.8g9ts8gm5). The "Learn to Code!" buzz is not the same everywhere in the world, but it is clear that there is a need for these skills across almost every domain. But who is trying to learn? And who is failing? How might we intervene? You are a learner too, and it will be fun to compare your own answers to the survey as we go along. Today's lesson will be a warmup in R, Statistics, and handling data. 

```{r}
library(ggplot2)
library(dplyr)

```

```{r}
data <- read.csv("../data/newcoders.csv")
head(data)

# how many people are in this dataset?
nrow(data)

#R doesn't have a built-in mode, but this means find the highest count for any of the values and return that value
getmode <- function(v) {
   uniqv <- unique(na.omit(v))
   uniqv[which.max(tabulate(match(v, uniqv)))]
}

#mean, median, and mode

mean_age <-mean(data$Age,na.rm=TRUE)
med_age <- median(data$Age,na.rm=TRUE)
mode_age <-getmode(data$Age)

ggplot(data,aes(Age))+
  geom_histogram()+
  geom_vline(xintercept = mean_age,linetype="dashed",color="red")+
  geom_vline(xintercept = med_age,linetype="dashed",color="blue")+
  geom_vline(xintercept = mode_age,linetype="dashed",color="green")+
  theme_bw()

median(data$Age,na.rm=TRUE)

table(data$Gender)

ggplot(data,aes(Gender,fill=Gender))+
  geom_bar()+
  theme_bw()
```

```{r}
library(dplyr)
forcats::fct_explicit_na

bootcampers <- data[!is.na(data$Gender)&!is.na(data$BootcampFinish),]%>%
  group_by(Gender) %>%
  summarise(mean=mean(BootcampFinish))
bootcampers

ggplot(bootcampers,aes(Gender,mean,fill=Gender))+
  geom_bar(stat="identity")+
  theme_bw()

table(data$Gender)

wilcox.test(data$BootcampFinish[data$Gender=='male'],data$BootcampFinish[data$Gender=='female'])
```