---
title: "Test-Driven Development: Does It Really Matter to Test-First or to Test-Last?"
author: "Yim Register"
date: "7/16/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


In my experience, the world of software is an *opinionated* one. It seems like everyone has special tools, techniques, practices, and packages that they swear by. In this lesson, we will investigate one of these classic claims: Which is better? Test-first or test-last development? We will also explore what it looks like when people misattribute benefits to a specific practice, when in reality the cause is entirely different. The study we are using is: [*A Dissection of the Test-Driven Development
Process: Does It Really Matter to Test-First or to
Test-Last?*](https://arxiv.org/pdf/1611.05994.pdf)
Below is an excellent quote from the paper, highlighting how statistics can separate the relevant findings from other distracting factors:

> *"This information can liberate developers and organizations
who are interested in adopting TDD, and trainers who
teach it, from process dogma based on pure intuition,
allowing them to focus on aspects that matter most in
terms of bottom line."*

## Test-Driven Development Process
- what is it?
- why do people care?

## The Study
- beyond their very complex design
- granularity, uniformity, sequencing, and refactoring
- test-first, test-last
```{r}
data <- read.csv("../bin/data/dissectionTDD/dataset.csv",sep=";")
head(data)
```

## Outcomes We Care About

> *Which subset of the factors best explain
the variability in external quality?*

> *Which subset of the factors best explain
the variability in developer productivity?*

QLTY, TUS
PROD


## Factors We Measure



## Descriptive Statistics
There is a difference between **descriptive statistics** and **inferential statistics**. Descriptive statistics describe properties of the data; such as **means**, **ranges**, and **normality** of the variables of interest. Inferential statistics will draw the actual *conclusions* about the data; reporting on **correlations**, **hypothesis tests**, and **estimation of parameters**. Inferential statistics helps to generalize about a larger **population**, that can go beyond the descriptive statistics of an immediate sample.

```{r}
library(ggplot2)

#small function to generate colors for ggplot
gg_color_hue <- function(n) {
  hues = seq(15, 375, length = n + 1)
  hcl(h = hues, l = 65, c = 100)[1:n]
}

titles <- c("Quality","Productivity","Granularity","Uniformity","Sequencing","Refactoring")
#get some colors for each
cols = gg_color_hue(length(titles))

# loop to create 6 density plots to look at spread for each variable
loop <- 2:7
for( i in loop){
  x <- data[[i]]
  plt <- ggplot(data,aes(x)) + 
    ggtitle(paste("Histogram and Density for",titles[i-1]))+
  geom_histogram(aes(y = ..density..), bins=25,color="black",fill=cols[i-1])+
  geom_density(aes(y = ..density..),color="black",fill="black", alpha=.2,stat = 'density')+
    xlab(titles[i-1])+
  theme_bw()
  
  print(plt)
  print(shapiro.test(x))
}
```

## Let's Talk About Interactions
- discuss interactions here
```{r}
cor(data$REF,data$SEQ)
## Component Residual Plots

```

## Component + Residual Plots
```{r}
library(car)

crPlots(m<-lm(QLTY~GRA+UNI+SEQ+REF,data=data))

```

## Regression Analysis vs. Hypothesis Testing
- explain how because theyre all continuous variables, we dont do hypothesis testing. we use regression