---
title: "Code Quality and Lines of Code"
author: "Yim Register"
date: "6/27/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Code Quality vs. Lines of Code

### Redundancy?
- Notes for me:
- Landman suggests not redundant enough to chuck the CC metric
- The correlation between aggregated CC for all subroutines and the total SLOC of a file is higher than the correlation between CC and SLOC of individual subroutines


## How Does Complexity Relate to Quality?










## Cycolmatic Complexity



## "Big Data"! Lines of Code vs. Function Declarations
We saw in our last exercise that we could use ASTs to explore the components of a program. That was a helpful exercise for learning about how we employ tools to collect data from programs on GitHub. However, we only collected a few hundred programs; and if you recall, they were all implementations of data structures and programming concepts in isolation. We know that from our *sample*, our results may not generalize to production code. This means we are lacking something called **external validity**: results of a study can realiably generalize to other contexts/samples (within reason). 

Now, you've probably heard this before when discussing concepts like Big O: it only really starts to matter for performance when n is big enough. I know I've certainly questioned why *O(log(n))* vs. *O(n)* even matters when my little sorting snippet runs just fine whatever the algorithm. But here is a big concept: **Complexity matters for big data, and the important findings may be undetectable on toy problems**.

Landman et al. also investigated this problem, using *millions* of Java and C programs. 


### theyre actually different problems. I have to figure out how to tie them together: our problem looks at functions and lines, theirs looks at number of lines per function, and also looks at CC
```{r}
# lang,cc,sloc

source("../bin/data/ESEUR_config.r")

library("plyr")
cc_loc=read.csv(paste0(ESEUR_dir, "../bin/data/Landman_m_ccsloc.csv.xz"), as.is=TRUE)
cc_loc=subset(cc_loc, cc != 0)
cc_loc$logloc=log(cc_loc$sloc)

C_loc=subset(cc_loc, lang == "C")
Java_loc=subset(cc_loc, lang == "Java")


```

